---
output:
  md_document:
    variant: markdown_github
---


# textsearch

A simple framework for searching corpora for lexical & grammatical patterns in context.  At present, functions facilitate two types of search: (1) TIF search, ie, raw, un-annotated text search, and (2) annotated corpus search.  


For dependency-based search, see [this lovely package]().

```{r include=FALSE}
library(tidyverse)
```



```{r eval=FALSE, include=FALSE}
With some add-on functionality for featurizing texts per [Biber (1988)](https://scholar.google.com/citations?view_op=view_citation&hl=en&user=mdWIU4MAAAAJ&alert_preview_top_rm=2&citation_for_view=mdWIU4MAAAAJ:u-x6o8ySG0sC).

*Biber blog post* -- 

https://linguisticswithacorpus.wordpress.com/2021/12/22/corpus-linguistics-is-for-text-lovers%EF%BF%BC/


*EXCELLENT Jack Grieve resources* ::

https://rpubs.com/jwgrieve/340342
https://aclanthology.org/W17-3001.pdf


> Need to create a list of MWE from Biber -- 

```



## Installation

You can download the development version from GitHub with:

```{r eval=FALSE}
remotes::install_github("jaytimm/textsearch")
```



## TIF search

### Build corpus

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```


```{r message=FALSE, warning=FALSE}
rss1 <- quicknews::qnews_build_rss(x = 'political ideology')
meta <- quicknews::qnews_strip_rss(rss1) 
news <- quicknews::qnews_extract_article(url = meta$link, cores = 7)
tif <- merge(news, meta)
tif$doc_id <- c(1:nrow(tif))

list(Date = tif$date[1],
     Source = tif$source[1],
     Title = tif$title[1],
     Article = strwrap(tif$text[1], 
                       width = 60)[1:5])
```



### Queries

```{r}
parts <- '\\bpart[a-z]*\\b'
pols <- 'political \\w+'
terms <- c('populism', 'political ideology')
```


```{r}
textsearch::find_lex(query = pols,
                     text = tif$text,
                     doc_id = tif$doc_id,
                     window = 15,
                     highlight = c('`', '`')) %>%
  slice(1:3) %>%
  knitr::kable(escape = F)
```

```{r}
set.seed(99)
textsearch::find_lex(query = parts,
                     text = tif$text,
                     doc_id = tif$doc_id,
                     window = 15,
                     highlight = c('`', '`')) %>%
  sample_n(3) %>%
  knitr::kable(escape = F)
```




## Annotated corpus search


### Annotate corpus 

```{r include=FALSE}
locald <- '/home/jtimm/pCloudDrive/GitHub/packages/biberizer/model'
```


```{r message=FALSE, warning=FALSE}
setwd(locald)
udmodel <- udpipe::udpipe_load_model('english-ewt-ud-2.3-181115.udpipe')

annotation <- tif %>% 
  text2df::tif2sentence() %>%
  text2df::tif2token() %>%
  text2df::token2annotation(model = udmodel)
```



### Build inline TIF

```{r}
inline_tif <- textsearch::build_inline(df = annotation, 
                                       form = 'token', 
                                       tag = 'xpos')

strwrap(inline_tif$text[1], width = 60)[1:5]
```


### Build inline query 

```{r}
search <- textsearch::translate_query(
  x = '(is|was) (ADV)* VBN by',
  mapping = textsearch::mapping_generic)

search
```




### Identify/extract grammatical constructions 

```{r}
found <- textsearch::find_gramx(tif = inline_tif, query = search)

found %>% slice(3:9) %>% knitr::kable()
```



### Add sentential context

```{r}
f_sentence <- textsearch::add_context(gramx = found,
                                      df = annotation,
                                      form = 'token', 
                                      tag = 'xpos',
                                      highlight = '`')

set.seed(99)
f_sentence %>% sample_n(5) %>% knitr::kable()
```




### Recode construction

```{r}
new_annotation <- textsearch::recode_gramx(df = annotation,
                                           gramx = found,
                                           form = 'token', 
                                           tag = 'xpos',
                                           
                                           col = 'xpos',
                                           new_cat = 'by_passive',
                                           renumber = T)
```



```{r eval=FALSE, include=FALSE}
#doc = 4; sentence = 23; df = new_annotation

see_example <- function(df,
                        doc, 
                        sentence){
  
  eg <- df[, if(any(doc_id %in% doc) & any(sentence_id %in% sentence)) .SD, 
                      by = list(doc_id, sentence_id)]
  
  eg0 <- eg[, list(text = paste(token, collapse = " ")), 
                                  by = list(doc_id,sentence_id)]
  
  eg0$text
  
  }
```



```{r message=FALSE, warning=FALSE}
new_annotation %>%
  group_by(doc_id, sentence_id) %>%
  filter(any(xpos == 'by_passive')) %>%
  mutate(token = ifelse(xpos == 'by_passive', 
                        paste0('`', token, '`'), 
                        token)) %>%
  summarize(example = paste0(token, collapse = ' ')) %>%
  ungroup() %>%
  sample_n(5) %>%
  knitr::kable()
```



## Summary
