---
output:
  md_document:
    variant: markdown_github
---


## linguistic features


```{r eval=FALSE, include=FALSE}
*Biber blog post* -- 

https://linguisticswithacorpus.wordpress.com/2021/12/22/corpus-linguistics-is-for-text-lovers%EF%BF%BC/


*EXCELLENT Jack Grieve resources* ::

https://rpubs.com/jwgrieve/340342
https://aclanthology.org/W17-3001.pdf


> Need to create a list of MWE from Biber -- 

```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```




## Build a corpus from current news

```{r eval=FALSE, message=FALSE, warning=FALSE}
rss1 <- lapply(c('economy', 
                 'covid', 
                 'biden', 
                 'depression', 
                 'merkel', 
                 'trump', 
                 'jobs'),
               
               quicknews::qnews_build_rss)

meta <- lapply(unlist(rss1), quicknews::qnews_strip_rss)
meta <- data.table::rbindlist(meta)
meta <- unique(meta)

news <- quicknews::qnews_extract_article(url = meta$link[1:100], 
                                         cores = 7) %>% 
  left_join(meta) %>%
  mutate(doc_id = row_number() )
```



```{r eval=FALSE, include=FALSE}
# setwd('/home/jtimm/Desktop')
# saveRDS(news, 'tif.rds')
```



## Annotate corpus 

```{r include=FALSE}
locald <- '/home/jtimm/pCloudDrive/GitHub/packages/biberizer/model'
```


```{r eval=FALSE, message=FALSE, warning=FALSE}
setwd(locald)
udmodel <- udpipe::udpipe_load_model('english-ewt-ud-2.3-181115.udpipe')

anno <- news %>% 
  text2df::tif2sentence() %>%
  text2df::tif2token() %>%
  #text2df::token2mwe(mwe = MWE) %>%
  text2df::token2annotation(model = udmodel)
```


```{r message=FALSE, warning=FALSE, include=FALSE}
# setwd('/home/jtimm/Desktop')
# saveRDS(anno, 'anno.rds')

setwd('/home/jtimm/Desktop')
anno <- readRDS('anno.rds')
tif <- readRDS('tif.rds')
```




## Search process -- 

### Build inline TIF


```{r}
inline_tif <- lxfeatures::build_inline(df = anno, 
                           form = 'token', 
                           tag = 'xpos')
```



### Optionally set tag mappings

```{r}
generic_mapping  <- list(V = c('VB', 'VBD', 'VBG', 'VBN', 'VBP', 'VBZ'),
                         N = c('NN', 'NNP', 'NNPS', 'NNS'),
                         ADJ = c('JJ', 'JJR', 'JJS'),
                         ADV = c('RB', 'RBR', 'RBS')
                         )
```



### Build inline query 

```{r}
## f18 = 'BE (ADV)* VBN by'
search <- lxfeatures::translate_query(x = 'VBG the ADJ N', mapping = generic_mapping)

search
```




### Identify - extract grammatical constructions -- 

```{r}
found <- lxfeatures::find_gramx(tif = inline_tif, query = search)

found %>% head() %>% knitr::kable()
```



### Add sentential context to a `gramx` object --

```{r}
f_sentence <- lxfeatures::add_context(gramx = found,
                          df = anno,
                          form = 'token', 
                          tag = 'xpos',
                          highlight = '`')

f_sentence %>% sample_n(5) %>% knitr::kable()
```




### Recode constructions -- per annotated DF --

```{r}
simple_np <- '(DT)?(ADJ|N)+(N)+'

search_np <- lxfeatures::translate_query(x = simple_np, mapping = generic_mapping)
```



```{r}
found0 <- lxfeatures::find_gramx(tif = inline_tif, query = search_np)

new_anno <- lxfeatures::recode_gramx(df = anno,
                         gramx = found0,
                         form = 'token', 
                         tag = 'xpos',
                         col = 'xpos',
                         new_cat = 'NPhrase',
                         renumber = T
                         )
```



## Some Biber (1988) odds/ends -- 

From: Biber, D. (1988). *Variation across speech and writing*. Cambridge University Press.

### Lexical features

```{r message=FALSE, warning=FALSE}
anno[, biber := lxfeatures::biber_tags(token = anno$token, tag = anno$xpos)]
```




### Lexico-grammatical features

```{r}
biber_mapping  <- list(V = c('VB', 'VBD', 'VBG', 'VBN', 'VBP', 'VBZ'),
                       N = c('NN', 'NNP', 'NNPS', 'NNS'),
                       ADJ = c('JJ', 'JJR', 'JJS'),
                       ADV = c('RB', 'RBR', 'RBS'),
                       AUX = c('MODAL', 'HAVE', 'BE', 'DO', 'S'),
                       PRO = c('SUBJPRO', 'OBJPRO', 
                               'POSSPRO', 'REFLEXPRO', 
                               'YOU', 'HER', 'IT'),
                       DET = c('ART', 'DEM', 'QUAN', 'NUM'),
                       ALLP = c('CLP', ','),
                       
                       ## may want to add to verb list -- no ??
                       PRV = c('SUAPRV', 'PRV'),
                       PUB = c('SUAPUB', 'PUB'),
                       SUA = c('SUA', 'SUAPUB', 'SUAPRV')
                       )
```
 
 

Some patterns:


```{r}
zz <- list(f18 = 'BE (ADV)* VBN by',
           f29 = 'NOUN that (ADV)* VBD', # *
           f19 = 'BE (DET|POSSPRO|PREP|ADJ)',
           f22 = 'ADJ that',
           f30 = 'NOUN that (DET|SUBJPRO|POSSPRO)', # *
           f33 = 'PREP WHP',
           f62 = 'to (ADV)+ VB',
           f61 = 'PREP CL-P',
           cc = 'MODAL V') # *
```

